HO unification from object language to meta language

### 2024-03-28 paper/main.tex
deduplicate-map: better formula

\newcommand{\cv}{\ensuremath{\mathrm{cv}}}
\newcommand{\ev}{\ensuremath{\mathrm{ev}}}
\newcommand{\ar}{\ensuremath{\mathrm{ar}}}

Before formally defining this procedure, we need to define some auxiliary
relations. Let \mapStore be the list of mapping and let $m \in \mapStore$, such
that $m = \mapping{A}{\textbf{A}}{n}$, we define $\cv, \ev, ar$ such that
$\cv(m) = A$, $\ev(m) = \textbf{A}$ and $\ar(m) = n$. Let $\langle m_1, m_2
\rangle \in \mapStore$ such that $\ar(m_1) < \ar(m_2)$ and $n = ar(m_1) -
ar(m_2) - 1$. We also let $A^i$ be a fresh \Ho variable. We define the
\elpiIn{make-eta-link} relation taking two mappings $\langle m_1, m_2 \rangle$
and returning  \linketa such that:

$$ 
  \forall i \in [0 \dots n], \left\{ 
    \begin{array}{ll}
      \linketaM{\phantom{\Gamma:x_0\dots x_{i-1}}}
              {\ev(m_1)}
              {\lambda x.A^0_x} & if~ i = 0 \\
      \linketaM{\Gamma:x_0\dots x_{i-1}}
               {A^i_{\Gamma}}
               {\lambda x.A^{i+1}_{\Gamma\cup x}} & if~ 0 < i < n\\
      \linketaM{\Gamma:x_0\dots x_{i}\phantom{_{-1}}}
               {A^{i-1}_{\Gamma}}
               {\lambda x.\ev(m_2)_{\Gamma\cup x}} & if~ i = n\\
    \end{array} 
  \right. 
$$

More concretely, we are saying that for any two mappings, we build as many
\linketa as the difference of the arities between the two mappings. This links
are constructed in such a way that the \Ho variable $v$ with lowest arity is
linked to a fresh variable eta-expended variable $A^0$ having the scope of $v$.
This variable $A^0$ is then linked to a fresh variable $A^{1}$ with same scope
of $A^0$ and so on. The last link is built between the $A^{n-1}$ (where $n$ is
the difference of arities between the two mappings) and the \Ho variable $u$
with higher arity in the two mappings being considered.

\begin{definition}[map-deduplication]
  Forall mappings $\langle m_1, m_2 \rangle \in \mapStore$, such that $\cv(m_1) =
  \cv(m_2)$, the list of \linketa $L$ is created thanks to
  \elpiIn{make-eta-link}$m_1~m_2~L$ and is added to \linkStore. Then $m_1$ is
  removed from \mapStore.
\end{definition}

>>>>

Before formally defining this procedure, we need to define some auxiliary
relations. Let \mapStore be the list of mapping and $\langle m_1, m_2 \rangle
\in \mapStore$ such that the arity of the \Ho variable in $m_1$ is smaller than
the one in $m_2$. Let $X$ (resp. $Y$) the \Ho variable of $m_1$ (resp. $m_2$)
and $n = ar(m_1) - ar(m_2)$. We also let $A^i$ be a fresh \Ho variable. We
define the \elpiIn{make-eta-link} relation taking two mappings $\langle m_1, m_2
\rangle$ and returning the following list of link \linketa: $\forall i \in [1..n]$,
{
\def\scopeA{\ensuremath{x_1\dots x_{i-1}}}
\def\scopeB{\ensuremath{x_1\dots x_{i}}}
%
$$ 
  \left\{ 
    \begin{array}{ll}
      \linketaM{\phantom{\scopeA}}{X}      {\lambda x.  A^1_x}         & if~ i = 1 \\
      \linketaM{\scopeA}{A^{i-1}_{\scopeA}}{\lambda x_i.A^i_{\scopeB}} & if~ 1 < i < n\\
      \linketaM{\scopeA}{A^{i-1}_{\scopeA}}{\lambda x_i.Y  _{\scopeB}} & if~ i = n
    \end{array} 
  \right. 
$$
}
\noindent
More concretely, we are saying that for any two mappings, we build as many
\linketa as the difference of the arities between the two mappings. This links
are constructed in such a way that the \Ho variable $v$ with lowest arity is
linked to a fresh variable eta-expended variable $A^1$ having the scope of $v$.
This variable $A^1$ is then linked to a an $\eta$-expanded fresh variable $A^2$
with same scope of $A^1$ and so on. The last link is built between the $A^{n-1}$
(where $n$ is the difference of arities between the two mappings) and the \Ho
variable $u$ with higher arity in the two mappings being considered.

\begin{definition}[map-deduplication]
  Forall mappings $\langle m_1, m_2 \rangle \in \mapStore$, sharing the same
  \Fo{} variable, the list of \linketa $L$ is created thanks to
  \elpiIn{make-eta-link}$m_1~m_2~L$ and is added to \linkStore. Then $m_1$ is
  removed from \mapStore.
\end{definition}
